I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-17 Paper/Paper_noProof.tex
update

Consider the simpler abstraction, called the {\em box abstraction} \cite{deeppoly}: from the bounds in the input, one compute by the linear step the bound on the next layer, apply the ReLU function, and so on. For instance, on the DNN example of Fig \ref{fig1}, $x_1,x_2 \in [-1,1]$, then $x_3,x_4 \in [-2,2]$, using ReLU $\hat{x}_3,\hat{x}_4 \in [0,2]$, and $x_5 \in [0,6]$ while $x_6 \in [1.5,3.5]$. 
While the bounds of $x_1, \ldots, x_6$ are exact (for all $\alpha$ 
in the range, there is an input $\vy$ such that $\val_{\vy}(x_i)=\alpha$), 
it is no more the case from the next layer (starting in $x_5,x_6$) because of potential 
dependencies between earlier neurons: $x_3$ reaches weight $2$ only when $x_1=x_2=2$,
in which case $x_4$ reaches weight $0$. Hence, it is impossible that $weight(x_5)$ reaches $6$, as it would require both $x_3$ and $x_4$ to have weight $2$.

A still extremely fast algorithm that corrects some of this issue is known as {\em DeepPoly} \cite{deeppoly}, also rediscovered as the {\em CROWN} algorithm \cite{crown}. Instead of 
hard bounds, for each neuron $n$ of layer $k$, two affine functions (with input at the previous layer $k-1$) are kept as lower bound and upper bound of the weight of the node.
For instance, $f_i \leq weight(\hat{x}_i) \leq g_i$ with 
$f_{3}(x_3)=f_4(x_4)=0$ and 
$g_3(x_3) = \frac{x_3+2}{2}$,
$g_4(x_4) = \frac{x_4+2}{2}$. 
With that, we get 
$x_5 \leq g_3(x_3) + 2 g_4(x_4) = \frac{x_3 + 2x_4 + 6}{2} = \frac{3x_1 - x_2 + 6}{2}\leq 5.5$.

While there is one optimal linear function for the upper bound $g_i(x_i)= \beta \frac{x_i-\alpha}{\beta-\alpha}$ for ReLU nodes, there are two choices for the lower bound: $f_i(x_i) = 0$ or $f_i(x_i)=x_i$.
In the DeepPoly algorithm, the choice is made by considering the range of an unstable neuron 
$x_i=[\alpha,\beta]$ with $\alpha < 0 < \beta$. 
If $|\alpha|\geq |\beta|$ then $0$ is used, while if $|\beta|>|\alpha|$ then $x_i$ (the identity function) is used. Let us call $\overline{\mbox{DeepPoly}}$ the DeepPoly algorithm such that 
$f_i(x_i) = 0$ is always used, and $f_i(x_i) = x_i$ is never used. 
It has a main advantage, as it subsumes the box abstraction.
Indeed, taking an interval $x_i \in [-0.2,5]$, it gives $ReLU(x_i) \in [0,5]$, 
while the normal DeepPoly algorithm would only allow to conclude that $ReLU(x_i) \in [-0.2,5]$.

@@

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $weight(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of exact $-x_2 \leq 1$), obtaining a bound $weight(x_5) \leq 5.5$ instead of $4$.

>>>>

Consider the simpler abstraction, called the {\em box abstraction} \cite{deeppoly}: from the bounds in the input, one compute by the linear step the bound on the next layer, apply the ReLU function, and so on. For instance, on the DNN example of Fig \ref{fig1}, $x_1,x_2 \in [-1,1]$, then $x_3,x_4 \in [-2,2]$, using ReLU $\hat{x}_3,\hat{x}_4 \in [0,2]$, and $x_5 \in [0,6]$ while $x_6 \in [0,2]$. 
While the bounds of $x_1, \ldots, x_4$ are exact (for all $\alpha$ 
in the range, there is an input $\vy$ such that $\val_{\vy}(x_i)=\alpha$), 
it is no more the case from the next layer (starting in $x_5,x_6$) because of potential 
dependencies between earlier neurons: $x_3$ reaches weight $2$ only when $x_1=x_2=2$,
in which case $x_4$ reaches weight $0$. Hence, it is impossible that $\Val(x_5)$ reaches $6$, as it would require both $x_3$ and $x_4$ to have weight $2$.

A still extremely fast algorithm that corrects some of this issue is known as {\em DeepPoly} \cite{deeppoly}, also rediscovered as the {\em CROWN} algorithm \cite{crown}. Instead of 
hard bounds, for each neuron $n$ of layer $k$, two affine functions (with input at the previous layer $k-1$) are kept as lower bound and upper bound of the weight of the node.
For instance, $f_i \leq \Val(\hat{x}_i) \leq g_i$ with 
$f_{3}(x_3)=f_4(x_4)=0$ and 
$g_3(x_3) = \frac{x_3+2}{2}$,
$g_4(x_4) = \frac{x_4+2}{2}$. 
With that, we get 
$x_5 \leq g_3(x_3) + 2 g_4(x_4) = \frac{x_3 + 2x_4 + 6}{2} = \frac{3x_1 - x_2 + 6}{2}\leq 5$.

While there is one optimal linear function for the upper bound $g_i(x_i)= \beta \frac{x_i-\alpha}{\beta-\alpha}$ for ReLU nodes, there are two choices for the lower bound: $f_i(x_i) = 0$ or $f_i(x_i)=x_i$.
In the DeepPoly algorithm, the choice is made by considering the range of an unstable neuron 
$x_i=[\alpha,\beta]$ with $\alpha < 0 < \beta$. 
If $|\alpha|\geq |\beta|$ then $0$ is used, while if $|\beta|>|\alpha|$ then $x_i$ (the identity function) is used. Let us call $\overline{\mbox{DeepPoly}}$ the DeepPoly algorithm such that 
$f_i(x_i) = 0$ is always used, and $f_i(x_i) = x_i$ is never used. 
It has a main advantage, as it subsumes the box abstraction.
Indeed, taking an interval $x_i \in [-0.2,5]$, it gives $\ReLU(x_i) \in [0,5]$, 
while the normal DeepPoly algorithm would only allow to conclude that $\ReLU(x_i) \in [-0.2,5]$.

@@

Compensating pair of paths will make precision of the target state reduce when using overabstraction of values. Here, $\Val(x_5) \leq 4$ (the value is reached for $x_1=1$ and $x_2=-1$). However, Box computes an upper bound of $6$, and while DeepPoly recovers some of the dependency in $x_2$, it can only cancel out some of $x_2$ value due to the ReLU over-abstraction (contribution of $\frac{-x_2+4}{2} \leq 2.5$ instead of exact $-x_2 \leq 1$), obtaining a bound $\Val(x_5) \leq 5$ instead of $4$.

### 2024-01-17 Paper/proofs.tex
update

@@

		Similarly, we can prove if $S(c)=-1$, $\min(c)=\Val_{\vx^*}(c)$. And similarly, we can prove this for $x^\sharp$. This is the induction step. 
	\end{proof}

@@

			Hence we have equality before and after ReLU in all cases.
		\end{proof}

@@

				Before the proof of above theorem, we will prove a weaker version, that is, consider not only compensating pairs, but also pairs that allows common nodes (the two paths in a compensating pairs cannot have a common nodes). We call such pair a weak compensating pair.

@@

					Second, we show that if neither 1 nor 2 hold, then $b$ will be on a weak compensating path.
					Because 2) does not hold, there is a $j \in J$ with $W_{a_j,b} \neq 0$, say $>0$.
					Because 1) does not hold, $W_{b,z} \neq 0$, say $>0$.
					Now, by definition of $J$, there is a pair of weak compensating paths $\pi,\pi'$ 
					from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
					Then $((a_j,b,z), \pi')$ is also a weak compensating pair.
				\end{proof}

@@

					3. if $b\in B_{zero}$, then $W_{bc}=0$; otherwise there exists $<b,c,z>$ a path from $b$ to $z$ the output node with a nonzero value, a contradiction.
				\end{proof}

@@

				\begin{proof}
					Otherwise, suppose there exists two paths $P_1,P_2$ from $b$ to $z$ with (strictly) different signs. Then consider $\langle a_i,b\rangle+P_1$ and $\langle a_i,b\rangle+P_2$, they are two paths with weights with different signs. This is a contradiction since $a_i\in I$.
				\end{proof}
				Therefore, for any node $b$  that $S$ is undefined on $b$, it only depends on input nodes in $J$.

@@

				\begin{proof}
					This is proved in the first section.
				\end{proof}

@@

				\begin{proof}
					This is because if there exists nodes $b,c$ such that $W_{bc}\neq 0$ and $c$ is in above set but $b$ is not, then $b$ has a path with nonzero weight from a node $a_i$ not in $J$ but this means so does $c$, a contradiction. Therefore any $b$ such that $W_{bc}\neq 0$ is also in above set.
				\end{proof}

@@

					For any fixed input vector $\boldsymbol{x}^0_J$, we can regard $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $\boldsymbol{x}^0_J$ %and their propagation in hidden layers 
					as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Lemma \ref{lem:key1} to get that $\Val_{x_I,\boldsymbol{x}^0_J}(z)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and we are done.
				\end{proof}

@@

				\begin{proof}
					For $D_I$, it is because of no compensating as proved in the first section. For $D_J$, it is because all nodes are either opened or constants.
				\end{proof}

@@

				\begin{proof}
					This is by the definition of MILP formulation and $\UB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}, \LB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}$.
				\end{proof}

@@

				\end{proof}

>>>>

	Before the proof of two theorems, we will prove a weaker version, that is, consider not only compensating pairs, but also pairs that allows common nodes (the two paths in a compensating pairs cannot have a common nodes). We may call such pair a weak compensating pair.

@@

		Similarly, we can prove if $S(c)=-1$, $\min(c)=\Val_{\vx^*}(c)$. And similarly, we can prove this for $x^\sharp$. This is the induction step. \qed
	\end{proof}

@@

			Hence we have equality before and after ReLU in all cases. \qed
		\end{proof}

		\subsubsection*{From weak to normal}

		\begin{lemma}
			If a DNN has no compensating pair, then it also does not have weak compensating pair.
		\end{lemma}

		\begin{proof}
			If it has a weak compensating pair of two path, $P_1,P_2$, then there exists a segment in two paths starting from the same node and end with the same node, but the weight along two segment have different signs. And this will also be a compensating pair. \qed
		\end{proof}

		This finishes the proof for compensating pair rather than weak version.

@@

@@

					Second, we show that if neither 1 nor 2 hold, then $b$ will be on a weak compensating path.
					Because 2) does not hold, there is a $j \in J$ with $W_{a_j,b} \neq 0$, say $>0$.
					Because 1) does not hold, $W_{b,z} \neq 0$, say $>0$.
					Now, by definition of $J$, there is a pair of weak compensating paths $\pi,\pi'$ 
					from $a_j$ to $z$, say with $\pi'$ with weight $<0$.
					Then $((a_j,b,z), \pi')$ is also a weak compensating pair. \qed
				\end{proof}

@@

					3. if $b\in B_{zero}$, then $W_{bc}=0$; otherwise there exists $<b,c,z>$ a path from $b$ to $z$ the output node with a nonzero value, a contradiction. \qed
				\end{proof}

@@

				\begin{proof}
					Otherwise, suppose there exists two paths $P_1,P_2$ from $b$ to $z$ with (strictly) different signs. Then consider $\langle a_i,b\rangle+P_1$ and $\langle a_i,b\rangle+P_2$, they are two paths with weights with different signs. This is a contradiction since $a_i\in I$. \qed
				\end{proof}
				Therefore, for any node $b$  that $S$ is undefined on $b$, it only depends on input nodes in $J$.

@@

				\begin{proof}
					This is proved in the first section. \qed
				\end{proof}

@@

				\begin{proof}
					This is because if there exists nodes $b,c$ such that $W_{bc}\neq 0$ and $c$ is in above set but $b$ is not, then $b$ has a path with nonzero weight from a node $a_i$ not in $J$ but this means so does $c$, a contradiction. Therefore any $b$ such that $W_{bc}\neq 0$ is also in above set. \qed
				\end{proof}

@@

					For any fixed input vector $\boldsymbol{x}^0_J$, we can regard $\Val_{\boldsymbol{x}_I,\boldsymbol{x}^0_J}(z)$ as a DNN $D'$ with input nodes $a_i\in I$, with all $\boldsymbol{x}^0_J$ %and their propagation in hidden layers 
					as bias. In the simplified $D'$, there is no compensating path because of the definition of $I$. Therefore we can apply Lemma \ref{lem:key1} to get that $\Val_{x_I,\boldsymbol{x}^0_J}(z)$ reaches its maximal value for $\boldsymbol{x}_I=\boldsymbol{x}_I^*$, and we are done. \qed
				\end{proof}

@@

				\begin{proof}
					For $D_I$, it is because of no compensating as proved in the first section. For $D_J$, it is because all nodes are either opened or constants. \qed
				\end{proof}

@@

				\begin{proof}
					This is by the definition of MILP formulation and $\UB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}, \LB_{\boldsymbol{x}^0_I,\boldsymbol{x}^0_J}$. \qed
				\end{proof}

@@

					\qed
				\end{proof}
